### 2024-12-25 machine-learning.qmd
20241025 - ensemble methods
20241225 - convert to calculate_stats() from nflfastR v5.0

# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))

@@

```{r}
# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = "./data/nfl_pbp.RData")
load(file = "./data/nfl_4thdown.RData")
load(file = "./data/nfl_participation.RData")
load(file = "./data/nfl_actualStats_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")

@@

@@

varNames <- c(
  names(nfl_players),
  names(nfl_teams),
  names(nfl_rosters),
  names(nfl_rosters_weekly),
  names(nfl_schedules),
  names(nfl_combine),
  names(nfl_draftPicks),
  names(nfl_depthCharts),
  names(nfl_pbp),
  names(nfl_4thdown),
  names(nfl_participation),
  names(nfl_actualStats_offense_weekly),
  names(nfl_injuries),
  names(nfl_snapCounts),
  names(nfl_espnQBR),
  names(nfl_nextGenStats_weekly),
  names(nfl_advancedStatsPFR_seasonal),
  names(nfl_advancedStatsPFR_weekly),
  names(nfl_playerContracts),
  names(nfl_ftnCharting),
  names(nfl_playerIDs),
  names(nfl_rankings),
  names(nfl_expectedFantasyPoints_weekly),
  names(nfl_expectedFantasyPoints_pbp)
)

@@

nfl_combine$pfr_id
nfl_advancedStatsPFR$pfr_id
nfl_advancedStatsPFR$pfr_player_id

@@

```{r}
# Prepare data for merging
#-todo: calculate years_of_experience
## Use common name for the same (gsis_id) ID variable
nfl_actualStats_offense_weekly <- nfl_actualStats_offense_weekly %>% 
  rename(gsis_id = player_id)

nfl_actualStats_offense_seasonal <- nfl_actualStats_offense_seasonal %>% 
  rename(gsis_id = player_id)

player_stats_seasonal_offense <- player_stats_seasonal_offense %>% 
  rename(gsis_id = player_id)

player_stats_weekly_offense <- player_stats_weekly_offense %>% 
  rename(gsis_id = player_id)

@@

```{r}
# Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
#-todo: remove redundant variables
playerListToMerge <- list(
  nfl_players %>% filter(!is.na(gsis_id)) %>% select(-season),
  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season)
)

@@

```{r}
seasonalData_lag <- seasonalData %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  mutate(
    fantasyPoints_lag = lead(fantasy_points)
  ) %>% 
  ungroup()

seasonalData_lag %>% 
  select(gsis_id, player_display_name, season, fantasy_points, fantasyPoints_lag)
```

@@

```{r}
seasonalData_lag_subsetQB <- seasonalData_lag_subset %>% 
  filter(position == "QB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    completions:rushing_2pt_conversions, special_teams_tds, contains(".pass"), contains(".rush"))

seasonalData_lag_subsetRB <- seasonalData_lag_subset %>% 
  filter(position == "RB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_lag_subsetWR <- seasonalData_lag_subset %>% 
  filter(position == "WR") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_lag_subsetTE <- seasonalData_lag_subset %>% 
  filter(position == "TE") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))
```

@@

newData_qb %>% 
  arrange(-fantasyPoints_lag) %>% 
  select(name, fantasyPoints_lag, fantasy_points)
```

>>>>

# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_pbp.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_4thdown.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_participation.RData", fsep = ""))
load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_weekly.RData", fsep = ""))
load(file = file.path(path, "/OneDrive - University of Iowa/Teaching/Courses/Fantasy Football/Data/nfl_expectedFantasyPoints_pbp.RData", fsep = ""))

@@

```{r}
# Downloaded Data - Processed
load(file = "./data/nfl_players.RData")
load(file = "./data/nfl_teams.RData")
load(file = "./data/nfl_rosters.RData")
load(file = "./data/nfl_rosters_weekly.RData")
load(file = "./data/nfl_schedules.RData")
load(file = "./data/nfl_combine.RData")
load(file = "./data/nfl_draftPicks.RData")
load(file = "./data/nfl_depthCharts.RData")
load(file = "./data/nfl_pbp.RData")
load(file = "./data/nfl_4thdown.RData")
load(file = "./data/nfl_participation.RData")
load(file = "./data/nfl_actualFantasyPoints_weekly.RData")
load(file = "./data/nfl_injuries.RData")
load(file = "./data/nfl_snapCounts.RData")
load(file = "./data/nfl_espnQBR_seasonal.RData")
load(file = "./data/nfl_espnQBR_weekly.RData")
load(file = "./data/nfl_nextGenStats_weekly.RData")
load(file = "./data/nfl_advancedStatsPFR_seasonal.RData")
load(file = "./data/nfl_advancedStatsPFR_weekly.RData")
load(file = "./data/nfl_playerContracts.RData")
load(file = "./data/nfl_ftnCharting.RData")
load(file = "./data/nfl_playerIDs.RData")
load(file = "./data/nfl_rankings_draft.RData")
load(file = "./data/nfl_rankings_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_weekly.RData")
load(file = "./data/nfl_expectedFantasyPoints_pbp.RData")

@@

*Ensemble* machine learning methods combine multiple machine learning approaches with the goal that combining multiple approaches might lead to more accurate predictions that any one method might be able to achieve on its own.

@@

varNames <- c(
  names(nfl_players),
  names(nfl_teams),
  names(nfl_rosters),
  names(nfl_rosters_weekly),
  names(nfl_schedules),
  names(nfl_combine),
  names(nfl_draftPicks),
  names(nfl_depthCharts),
  names(nfl_pbp),
  names(nfl_4thdown),
  names(nfl_participation),
  names(nfl_actualFantasyPoints_player_weekly),
  names(nfl_injuries),
  names(nfl_snapCounts),
  names(nfl_espnQBR_seasonal),
  names(nfl_espnQBR_weekly),
  names(nfl_nextGenStats_weekly),
  names(nfl_advancedStatsPFR_seasonal),
  names(nfl_advancedStatsPFR_weekly),
  names(nfl_playerContracts),
  names(nfl_ftnCharting),
  names(nfl_playerIDs),
  names(nfl_rankings_draft),
  names(nfl_rankings_weekly),
  names(nfl_expectedFantasyPoints_weekly),
  names(nfl_expectedFantasyPoints_pbp)
)

@@

nfl_combine$pfr_id
nfl_advancedStatsPFR_seasonal$pfr_id
#nfl_advancedStatsPFR_seasonal$pfr_player_id

@@

```{r}
# Prepare data for merging
#-todo: calculate years_of_experience
## Use common name for the same (gsis_id) ID variable
nfl_actualFantasyPoints_player_weekly <- nfl_actualFantasyPoints_player_weekly %>% 
  rename(gsis_id = player_id)

nfl_actualFantasyPoints_player_seasonal <- nfl_actualFantasyPoints_player_seasonal %>% 
  rename(gsis_id = player_id)

player_stats_seasonal_offense <- player_stats_seasonal %>% 
  filter(position_group %in% c("QB","RB","WR","TE")) %>% 
  rename(gsis_id = player_id)

player_stats_weekly_offense <- player_stats_weekly %>% 
  filter(position_group %in% c("QB","RB","WR","TE")) %>% 
  rename(gsis_id = player_id)

@@

```{r}
# Create lists of objects to merge, depending on data structure: id; or id-season; or id-season-week
#-todo: remove redundant variables
playerListToMerge <- list(
  nfl_players %>% filter(!is.na(gsis_id)),
  nfl_draftPicks %>% filter(!is.na(gsis_id)) %>% select(-season)
)

@@

```{r}
seasonalData_lag <- seasonalData %>% 
  arrange(gsis_id, season) %>% 
  group_by(gsis_id) %>% 
  mutate(
    fantasyPoints_lag = lead(fantasyPoints)
  ) %>% 
  ungroup()

seasonalData_lag %>% 
  select(gsis_id, player_display_name, season, fantasyPoints, fantasyPoints_lag)
```

@@

```{r}
seasonalData_lag_subsetQB <- seasonalData_lag_subset %>% 
  filter(position == "QB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    completions:rushing_2pt_conversions, special_teams_tds, contains(".pass"), contains(".rush"))

seasonalData_lag_subsetRB <- seasonalData_lag_subset %>% 
  filter(position == "RB") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_lag_subsetWR <- seasonalData_lag_subset %>% 
  filter(position == "WR") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))

seasonalData_lag_subsetTE <- seasonalData_lag_subset %>% 
  filter(position == "TE") %>% 
  select(
    gsis_id, season, games, gs, years_of_experience, age, ageCentered20, ageCentered20Quadratic,
    height, weight, rookie_year, draft_number,
    fantasy_points, fantasy_points_ppr, fantasyPoints, fantasyPoints_lag,
    carries:special_teams_tds, contains(".rush"), contains(".rec"))
```

@@

newData_qb %>% 
  arrange(-fantasyPoints_lag) %>% 
  select(name, fantasyPoints_lag, fantasyPoints)
```
